---
title: "![R](R) UNICEF"
author: "Prasanth Yadav Ch"
date: "04/27/2023"
format: html
execute: 
  echo: false
  warning: false
  message: false
---
</style>

 <style>
 body{
 text-align: justify}
 </style>

## What proportion of schools has basic sanitation?
 
- The **world map chart** shows the average proportion of schools with basic sanitation across different countries. Upon further exploration, the **scatter plot** shows that countries with greater average GDP per capita have a better proportion of schools with basic sanitation than countries with lower average GDP per capita. 

- When analyzed further, from the **time series chart**, it can be found that the average life expectancy and the average proportion of schools with basic sanitation are positively correlated because a lack of basic sanitation can increase the transmission of diseases. When the average life expectancy at birth and the average proportion of schools with basic sanitation is observed across the years, as the proportion of schools with basic sanitation decreased, the average life expectancy also reduced. 

- Lastly, the **bar chart** highlights the disparities in the proportion of schools with basic sanitation across different continents.  

Therefore, ***it is high time that the public should focus on improving basic sanitation in schools to achieve greater life expectancy***.

## World map chart

In the below world map, *the lighter the blue color's intensity higher the proportion of schools with basic sanitation*. The legend name *value indicates the avg proportion of schools with basic sanitation*  
 
```{r}
#| echo: false
#| include: false
library(ggplot2)
library(dplyr)
library(maps)
png("world_map.png", width=30000, height=600) # Set width to 1200 pixels
map("world")
dev.off() # Close the output device

 
# Load data
indicator1 <- read.csv("unicef_indicator_1.csv")
 
 
grouped_data <- indicator1 %>%
  group_by(country) %>%
  summarize(value = mean(obs_value))
 
world_map <- map_data("world")
 
map_res <- full_join(world_map, grouped_data, by = c("region" = "country"))
 
res <- ggplot(data = map_res) +
  aes(x = long, y = lat, group = group, fill = value) +
  geom_polygon()

```
 
```{r}
#| echo: false
res
```
 
## Scatter Plot with regression line 

The scatterplot identifies that *no country with above 40,000$ GDP per capita has a value of less than 100 for the average proportion of schools with basic sanitation*. 

 
```{r}
#| echo: false
#| include: false
library(ggplot2)
library(dplyr)
library(maps)
 install.packages("countrycode") # Install the package
library(countrycode) # Load the package

# Load data
indicator1 <- read.csv("unicef_indicator_1.csv")
metadata <- read.csv('unicef_metadata.csv')
 
colnames(metadata)[8] <- 'gdp_val'
 
metadata <- metadata %>%
  filter(gdp_val != "NA")

sanitization_df <- indicator1 %>% 
  group_by(country) %>% 
  summarize(avg_proportion_of_schools_with_basic_sanitation = mean(obs_value))

gdp_df <- metadata %>% 
  group_by(country) %>% 
  summarize(avg_gdp_per_capita = mean(gdp_val))



res <- full_join(sanitization_df, gdp_df, by = c("country"))

res <- res %>%
  filter(avg_gdp_per_capita != "NA")%>%
  filter(avg_proportion_of_schools_with_basic_sanitation != "NA")

res$continent <- countrycode(sourcevar = res$country, origin = "country.name", destination = "continent")


result <- ggplot(res)+
  aes(x = avg_proportion_of_schools_with_basic_sanitation, y = avg_gdp_per_capita,group ="continent")+
  geom_point(aes(color = continent)) +
  geom_smooth(method = "lm", se = FALSE)+
  #geom_point()+
  #geom_smooth(method = "lm", se = FALSE)+
  theme_bw()

 
```
 
```{r}
#| echo: false
result
```
## Time Series Chart

The time series chart below identifies the *positive correlation* between the avg life expectancy and avg proportion of schools with basic sanitation. Here *Value 1 is avg proportion of schools with basic sanitation and value 2 is avg life expectancy* 

 
```{r}
#| echo: false
#| include: false
library(ggplot2)
library(dplyr)
library(maps)
 
# Load data
indicator1 <- read.csv("unicef_indicator_1.csv")
metadata <- read.csv('unicef_metadata.csv')
 
metadata <- read.csv('Time series chart.csv')
 
 
 
colnames(metadata)[1] <- 'life_expectancy'
colnames(metadata)[3] <- 'avg_proportion_of_schools_with_basic_sanitation'
 
 
res <- ggplot(metadata, aes(x = year)) +
  geom_line(aes(y = avg_proportion_of_schools_with_basic_sanitation, color = "Value 1")) +
  geom_line(aes(y = life_expectancy, color = "Value 2")) +
  labs(title = "Line Charts of Two Variables Over Time", x = "Year", y = "Value") +
  ylim(0, 200) +
  theme_bw() +
  scale_color_manual(name = "", values = c("Value 1" = "blue", "Value 2" = "red"))
 
```
 
```{r}
#| echo: false
res
```
## Bar chart

From the below bar chart, it can be understood that the *most developed continents have a higher proportion of schools with basic sanitation*.   

 
```{r}
#| echo: false
#| include: false
library(ggplot2)
library(dplyr)
library(maps)
library(countrycode)
 
# Load data
indicator1 <- read.csv("unicef_indicator_1.csv")
colnames(indicator1)
 
indicator1$continent <- countrycode(sourcevar = indicator1$country, origin = "country.name", destination = "continent")
 
grouped_data <- indicator1 %>%
  group_by(continent) %>%
  summarize(avg_proportion_of_schools_with_basic_sanitation = mean(obs_value))
 
res <- grouped_data %>%
  ggplot(aes(continent, avg_proportion_of_schools_with_basic_sanitation, fill = continent)) +
  geom_col()
 
```
 
```{r}
#| echo: false
res
```
 
## Conclusion: 
To sum up, this report uses

1. World map chart

2. Scatter plot with regression line

3. Time series chart

4. Bar chart

***to raise awareness among the public on the world issue: proportion of schools with basic sanitation***.
   
 